---
title: "Chelsea's Quarto"
author: "Chelsea Sanford"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(sf)
library(terra)
library(tidyterra)
library(dplyr)
```

### Load in the Data

```{r}
ca_counties_raw_sf <- read_sf(here("data/ca_counties/CA_Counties_TIGER2016.shp"))
socioeco_data <- read.csv(here('data/socioeco.csv'))
water_quality <- read_csv(here('data/water_quality.csv'))
```

############################################## CES + County Data

### Edit the CES data

```{r}
socio_by_county <- socioeco_data %>% 
  dplyr::group_by(california_county) %>% 
  summarize(total_pop = sum(total_population, na.rm = TRUE),
            ces = mean(ces_4_0_percentile, na.rm = TRUE),
            low_birth_weight = mean(low_birth_weight_pctl, na.rm = TRUE),
            cardio_disease = mean(cardiovascular_disease_pctl, na.rm = TRUE),
            education = mean(education_pctl, na.rm = TRUE),
            poverty = mean(poverty_pctl, na.rm = TRUE),
            unemployment = mean(unemployment_pctl, na.rm = TRUE))

socio_by_county$county <- socio_by_county$california_county 

socio_county <- socio_by_county %>% 
  select(-california_county) %>% 
  select(county, everything()) %>% 
  mutate(county = str_squish(county))
```

### Create CA County Map and Overlay CES Score Data

What counties have the highest CES score?

```{r}
### counties data
ca_counties_sf <- ca_counties_raw_sf %>% 
  janitor::clean_names() %>%
  mutate(land_km2 = aland / 1e6) %>%
  select(county = name, land_km2) %>% 
  mutate(county = str_squish(county)) 
#ca_counties_sf %>% st_crs() ###3857

### join the two together
county_socio_join <- merge(x = ca_counties_sf, y = socio_county, by = "county", all.x = TRUE) %>% 
  mutate(density = total_pop/land_km2) %>% 
  select('county','density', everything(), -'total_pop', -'land_km2') %>% 
  pivot_longer('density':'unemployment', names_to = 'parameter', values_to = 'percentile')

### ggplot
pop_plot <- ggplot(data = county_socio_join) +
  geom_sf(aes(fill = density, geometry = geometry), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray", "orange","red")) +
  theme_void() +
  labs(fill = "Population Density")

plot(pop_plot)
```

######################################## Water Quality + CES

### Groundwater Quality vs Poverty Score

Questions we want to ask: Is the poverty score a good indicator of water quality? compare using NLS

```{r}
# # convert water quality to sf 
# water_quality_sf <- st_as_sf(x = water_quality, 
#                              coords = c('longitude', 'latitude'), 
#                              crs = 4326)
# 
# water_quality_sf <- st_transform(water_quality_sf, 3857)
# 
# 
# # combine water quality and county data 
# county_water <- st_join(ca_counties_sf, water_quality_sf)
# 
# county_water_year <- county_water %>%
#   drop_na() %>% 
#   separate(date, c("year", "month", "day")) %>% 
#   select(-month, -day, -land_km2) 
# 
# county_water_pivot <- county_water_year %>% 
#   group_by(year, county, chemical) %>%
#   summarize(mean_measurement = mean(measurement, na.rm = TRUE)) %>%
#   pivot_wider(names_from = chemical, values_from = mean_measurement) %>% 
#   janitor::clean_names()

# co3_plot <- ggplot(data = county_water_pivot) +
#   geom_sf(aes(fill = alkalinity_as_ca_co3, geometry = geometry), color = "white", size = 0.1) +
#   scale_fill_gradientn(colors = c("lightgray","orange","red")) +
#   theme_void() +
#   labs(fill = "Average CO3")
# 
# co3_plot
```

######################################### ShinyApp

### Attach necessary packages

```{r}
library(shiny)
library(tidyverse)
library(bslib)
```

### Create the user interface:

```{r}
vector <- c('total_pop', 'ces', 'low_birth_weight', 'cardio_disease', 'education', 'poverty', 'unemplpoyment')

ui <- fluidPage(
  theme = bs_theme(bootswatch = 'minty'),
  
  titlePanel('California Groundwater'),
  tabsetPanel(
    
    
    tabPanel( ### start tab 4
      title = 'Environmental Justice',
  
      fluidRow( # start fluid row 4.1
      column(width = 3,
                    h3('select demographic'),
                    radioButtons(
                      inputId = 'factor_4_1',
                      label = 'demographic category',
                      choices = c('Population Density', 'CES Score', 
                                  'Low Birth Weight', 'Cardiovascular Disease', 
                                  'Education', 'Poverty', 'Unemplpoyment')
                    )
        ), ### end column
        column(width = 9,
                    h3('Map Here'),
                    plotOutput(outputId = 'ces_plot')
        )
      ), ### end fluidRow 4.1
      
      hr(), ### horizontal rule so the row breaks are easier to see
      p('talk about how they should look at the map'),
      hr(),
      
      fluidRow( # start fluid row 4.2
        column(
          width = 3,
          h3('Select other variable'),
          radioButtons(
            inputId = 'demographic',
            label = 'demographic category',
            choices = c(4, 6, 8)
          )
        ),
        column(
          width = 9,
          h3('Graph here'),
          #plotOutput('insert_graph'))
        )
      ) ### end fluidRow 4.2
      
    ) ### end tab 4 
  
  )
) # end Fluidpage




```

### Create the server function:

```{r}
server <- function(input, output) {
  
### START tab 4, row 1
  soc_select <- reactive({ ### start soc_select
    soc_df <- county_socio_join %>% 
      filter(parameter == input$factor_4_1)
    #select(geometry, tmp = all_of(input$demographics))
    
    return(soc_df)
  }) ### end soc_select
  
    output$ces_plot <- renderPlot({ ### start ces_plot
       ggplot(data = soc_select) +
        geom_sf(aes(fill = soc_df, geometry = geometry), color = "white", size = 0.1) +
        scale_fill_gradientn(colors = c("lightgray", "orange","red")) +
        theme_void() +
        labs(fill = "Population") 

  
 }) ### end ces_plot
### END tab 4, row 1
 
}
```

### Combine them into an app:

```{r}
shinyApp(ui = ui, server = server)
```

### Home Page

### Groundwater depth levels

### Chemicals

### Socioeconomic

### groundwater dependent ecosystems ?
